Implicit yaddayadda

### 2011-10-26 papers/PLDI2012/implicitization.tex
rephrase

\citet{Monniaux_Gonnord_SAS11}'s \emph{path-focused} technique
distinguishes the different paths in the program in order to avoid loss of
precision due to merge operations. Since the number of paths may be exponential,
the technique keeps them implicit and computes them when needed using
SMT-solving.
The (accelerated) Kleene iterations (\S\ref{sec:static_analysis}) are computed over a reduced multigraph instead of the classical transition graph.

@@

We first check which paths become feasible, so that we can include them into the
next subset of paths. The set $P'$ of new feasible paths is added to the
previous set $P$. Each starting points of paths in $P'$ is added into $A$.
We apply the Path-focused technique\cite{Monniaux_Gonnord_SAS11} to
the elements of $A$ until $A$ is empty. 
When $A$ becomes empty, it means the
invariant computation of the subgraph is finished. As proposed by
\citet{DBLP:conf/sas/GopanR07}, we can do some narrowing iterations.
These narrowing iterations allow to recover precision lost by
widening, before computing and taking into account new feasible paths.
Thus, our technique combines both the advantages of \emph{Guided Static
Analysis} and \emph{Path-focusing}.

@@

The algorithm for computing an inductive invariant over a subgraph is the
Path-focused algorithm from \citet{Monniaux_Gonnord_SAS11} with special
treatment for self loops (line~\ref{alg=pf} in algorithm~\ref{algo:combined}).

@@

We distinguish two phases when analyzing a subgraph $P$:
\begin{enumerate}
\item We search for paths starting at point $p_i \in P_R$, such that these paths
	are in $P$, i.e are included in the working subgraph. Each time we find a
	path, we update the abstract value of the destination point of the path.
	This is the phase explained in \ref{subsec:ascending}.
\item  When the previous phase terminates, there is no more paths in
$P$, starting at $p_i$, 
that make the abstract values of the successors grow (otherwise, the SMT
solver would not have answered ``\emph{unsat}''). Narrowing iterations preserves
this property. 
However, there may exist such paths in the entire multigraph, that are not in
$P$. The second phase
computes these paths and adds them to $P'$ (line~\ref{alg=computeNewPaths} in
algorithm~\ref{algo:combined}). We also use SMT solving to discover
these new paths, but we slightly change the SMT formula given to the SMT solver:
we now simply check for the satisfiability of $f(p_i)$ instead of $f(p_i) \wedge
P$. Since we already know that $f(p_i) \wedge P$ is \emph{unsatisfiable}, all
the paths given by the SMT solver won't be in $P$.
\end{enumerate}

@@

\subsection{Termination}
Termination of this algorithm is guaranteed, because:
\begin{itemize}
\item 
the subset of paths $P$ is
strictly increasing at each loop iteration, and bounded by the finite set of
paths in the entire graph. 
\item the set $P'$ always verifies $P \cap P' = \emptyset$ by construction, 
which guarantees that $P'$
will eventually be empty after a finite number of loop iterations.
\end{itemize}

@@

We revise the rate limiter described in \ref{subsec:rate_lim}. In this example,
\emph{Path-focused} technique works well because all the paths starting at the
loop header are actually self loops. In such a case, the technique performs a
widening/narrowing sequence or accelerates the loop, thus leading to a precise
invariant. However, in real cases, there also exist paths that are not
self loops, where \emph{Path-focusing} only performs a widening. 
This widening may induce unrecoverable loss of precision.

@@

Guided static analysis from \citet{DBLP:conf/sas/GopanR07} yields at line 5
$\lstinline|x_old| \in~]-\infty,+\infty[$.
Path-focused technique from \citet{Monniaux_Gonnord_SAS11} performs slightly
better, and finds $\lstinline|x_old| \in~]-\infty,10000]$.

@@

This technique is aimed at computing for a loop header $p_i$ a disjunctive invariant
$\bigvee_{1\leq j \leq m} X_{i,j}$. 
To do so, one chooses an integer $\delta \in [1,m]$, and
a mapping function $\sigma: [1,m] \times [1,n] \mapsto [1,m]$, assuming there
are $n$ distinct paths in the loop.  For each
abstract value of the disjunctive invariant, and for each path $\tau_{i,k}$ in the loop, the
image of $X_{i,j}$ by the path is joined with
$X_{i,\sigma(j,k)}$.
Initially, the $\delta$-th abstract value is assigned to the initial states of
$p_i$, and each other abstract value is assigned to bottom.

@@

Our method improves this technique in two ways :
\begin{itemize}
\item Instead of enumerating the whole set of paths, we keep them implicit and
compute them only when needed.
\item At each loop iteration of the algorithm explained in
\cite{DBLP:conf/pldi/GulwaniZ10}, an image by a path and a union of abstract
values are computed for each path and each disjunct. 
Lots of these operations may be useless, in the sense that the resulting
abstract value may not change.
In our approach, we do this only for paths and disjuncts that makes the
resulting abstract value grow.
\end{itemize}

>>>>

\citet{Monniaux_Gonnord_SAS11}'s \emph{path-focusing} technique
distinguishes the different paths in the program in order to avoid loss of
precision due to merge operations. Since the number of paths may be exponential,
the technique keeps them implicit and computes them when needed using
SMT-solving.
The (accelerated) Kleene iterations (\S\ref{sec:static_analysis}) are computed over a reduced multigraph instead of the classical transition graph.

@@

We first check which paths become feasible, so that we can include them into the
next subset of paths. The set $P'$ of new feasible paths is added to the
previous set $P$. Each starting points of paths in $P'$ is added into $A$.
We apply the path-focusing \cite{Monniaux_Gonnord_SAS11} to
the elements of $A$ until $A$ is empty. 
When $A$ becomes empty, it means the
invariant computation of the subgraph is finished. As proposed by
\citet{DBLP:conf/sas/GopanR07}, we can do some narrowing iterations.
These narrowing iterations allow to recover precision lost by
widening, before computing and taking into account new feasible paths.
Thus, our technique combines both the advantages of \emph{Guided Static
Analysis} and \emph{Path-focusing}.

@@

For computing an inductive invariant over a subgraph, we use the
Path-focusing algorithm from \citet{Monniaux_Gonnord_SAS11} with special
treatment for self loops (line~\ref{alg=pf} in algorithm~\ref{algo:combined}).

@@

We distinguish two phases when analyzing a subgraph $P$:
\begin{enumerate}
\item We search for paths starting at point $p_i \in P_R$, such that these paths
	are in $P$, i.e are included in the working subgraph. Each time we find a
	path, we update the abstract value of the destination point of the path.
	This is the phase explained in \ref{subsec:ascending}.
\item  When the previous phase terminates, there are no more paths in
$P$, starting at $p_i$, 
that make the abstract values of the successors grow (otherwise, the SMT
solver would not have answered ``\emph{unsat}''). Narrowing iterations preserve
this property. 
However, there may exist such paths in the entire multigraph, that are not in
$P$. The second phase
computes these paths and adds them to $P'$ (line~\ref{alg=computeNewPaths} in
algorithm~\ref{algo:combined}). We also use SMT solving to discover
these new paths, but we slightly change the SMT formula given to the SMT solver:
we now simply check for the satisfiability of $f(p_i)$ instead of $f(p_i) \wedge
P$. Since we already know that $f(p_i) \wedge P$ is unsatisfiable, all
the paths given by the SMT solver won't be in $P$.
\end{enumerate}

@@

\subsection{Termination}
Termination of this algorithm is guaranteed, because:
\begin{itemize}
\item 
the subset of paths $P$ strictly increases at each loop iteration, and bounded by the finite set of
paths in the entire graph. 
\item the set $P'$ always verifies $P \cap P' = \emptyset$ by construction, 
which guarantees that $P'$
will eventually be empty after a finite number of loop iterations.
\end{itemize}

@@

We revise the rate limiter described in \ref{subsec:rate_lim}. In this example,
\emph{Path-focusing} works well because all the paths starting at the
loop header are actually self loops. In such a case, the technique performs a
widening/narrowing sequence or accelerates the loop, thus leading to a precise
invariant. However, in some cases, there also exist paths that are not
self loops, in which case \emph{Path-focusing} applies widening. 
This widening may induce unrecoverable loss of precision.

@@

Guided static analysis from \citet{DBLP:conf/sas/GopanR07} yields at line 5
$\lstinline|x_old| \in~]-\infty,+\infty[$.
Path-focusing \citep{Monniaux_Gonnord_SAS11} performs slightly
better, and finds $\lstinline|x_old| \in~]-\infty,10000]$.

@@

For a loop header $p_i$, this technique computes a disjunctive invariant
$\bigvee_{1\leq j \leq m} X_{i,j}$. 
To do so, one chooses an integer $\delta \in [1,m]$, and
a mapping function $\sigma: [1,m] \times [1,n] \mapsto [1,m]$, assuming there
are $n$ distinct paths in the loop.  For each
abstract value of the disjunctive invariant, and for each path $\tau_{i,k}$ in the loop, the
image of $X_{i,j}$ by the path is joined with
$X_{i,\sigma(j,k)}$.
Initially, the $\delta$-th abstract value is assigned to the initial states of
$p_i$, and each other abstract value is assigned to bottom.

@@

Our method improves this technique in two ways :
\begin{itemize}
\item Instead of enumerating the whole set of paths, we keep them implicit and
compute them only when needed.

\item At each loop iteration of the original algorithm \citep{DBLP:conf/pldi/GulwaniZ10}, an image by each path inside the loop is computed for each disjunct of the invariant candidate.
Yet, many of these images may be redundant: for instance, if our invariant candidate is $(0 \leq x \leq 10 \land 0 \leq y \leq 1000) \lor (x < -10 \land y < -10)$, then there is no point enumerating paths whose image is included in this invariant candidate.
In our approach, we compute such an image only if it makes the resulting abstract value grow.
\end{itemize}
